---
title: "Exp 2 — Clustering Univariate Gaussians, k=2"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
set.seed(42)
source("../../exp1_distances/R/distances.R")
source("../R/data_generation.R")
source("../R/clustering_k2.R")
dir.create("../results/tables",  recursive = TRUE, showWarnings = FALSE)
dir.create("../results/figures", recursive = TRUE, showWarnings = FALSE)
```

Replicates Zhang (2017) Table 8.1: K-Means and hierarchical clustering on
univariate Gaussian parameter estimates, comparing Fisher-Rao vs Euclidean
distance. Each algorithm is evaluated over 100 Monte Carlo replications.

---

## Parameters (Zhang Table 8.1)

```{r params}
mu_list    <- list(1, 2)
sigma_list <- list(1.5, 1.5)
k     <- 2L
t     <- 100L   # distributions per cluster
n     <- 30L    # samples per distribution
n_rep <- 100L   # Monte Carlo replications
```

True cluster parameters: $(\mu_1, \sigma_1) = (1, 1.5)$ and
$(\mu_2, \sigma_2) = (2, 1.5)$. Each replication generates $k \times t = 200$
estimated parameter pairs as the objects to cluster.

---

## Monte Carlo replications

```{r run, cache=TRUE}
results <- run_replications(
  mu_list, sigma_list,
  k = k, t = t, n = n, n_rep = n_rep
)
```

---

## Accuracy table

```{r summary}
algorithms <- c(
  "K-Means + Fisher-Rao",
  "K-Means + Euclidean",
  "Hierarchical + Fisher-Rao",
  "Hierarchical + Euclidean"
)
cols <- c("km_fr", "km_eu", "hc_fr", "hc_eu")

summary_df <- data.frame(
  Algorithm = algorithms,
  Accuracy  = sapply(cols, function(col) mean(results[[col]])),
  SE        = sapply(cols, function(col) sd(results[[col]]) / sqrt(n_rep)),
  row.names = NULL
)
summary_df$Replicated <- sprintf("%.3f ± %.3f", summary_df$Accuracy, summary_df$SE)

zhang_ref <- c(
  "0.965 ± 0.001",
  "0.965 ± 0.010",
  "0.904 ± 0.006",
  "0.922 ± 0.006"
)

comparison <- data.frame(
  Algorithm  = summary_df$Algorithm,
  Replicated = summary_df$Replicated,
  Zhang_2017 = zhang_ref
)

knitr::kable(
  comparison,
  col.names = c("Algorithm", "Replicated (mean ± SE)", "Zhang (2017) Table 8.1"),
  caption   = "Clustering accuracy over 100 Monte Carlo replications"
)
```

---

## Save results

```{r save}
write.csv(results,    "../results/tables/replication_results.csv", row.names = FALSE)
write.csv(summary_df, "../results/tables/accuracy_summary.csv",    row.names = FALSE)
cat("Results written to results/tables/\n")
cat("All checks passed:", all(comparison$Replicated != "NA ± NA"), "\n")
```
