---
title: "Exp 3 — Clustering Univariate Gaussians, k=3"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
set.seed(42)
source("../../exp1_distances/R/distances.R")
source("../R/data_generation.R")
source("../R/clustering_k3.R")
dir.create("../results/tables",  recursive = TRUE, showWarnings = FALSE)
dir.create("../results/figures", recursive = TRUE, showWarnings = FALSE)
```

Extends Zhang (2017) Ch.8 to k=3 univariate Gaussian clusters. The thesis does
not provide exact k=3 univariate parameters, so three well-separated distributions
are chosen with varying spread to expose Fisher-Rao's advantage over Euclidean
distance. Each algorithm is evaluated over 100 Monte Carlo replications.

---

## Parameters

```{r params}
mu_list    <- list(1, 2, 3)
sigma_list <- list(1.0, 1.5, 2.0)
k     <- 3L
t     <- 100L   # distributions per cluster
n     <- 30L    # samples per distribution
n_rep <- 100L   # Monte Carlo replications
```

True cluster parameters: $(\mu_1, \sigma_1) = (1, 1.0)$,
$(\mu_2, \sigma_2) = (2, 1.5)$, $(\mu_3, \sigma_3) = (3, 2.0)$.
Each replication generates $k \times t = 300$ estimated parameter pairs.
Varying $\sigma$ values make the Fisher-Rao metric non-trivially different
from Euclidean.

---

## Monte Carlo replications

```{r run, cache=TRUE}
results <- run_replications(
  mu_list, sigma_list,
  k = k, t = t, n = n, n_rep = n_rep
)
```

---

## Accuracy table

```{r summary}
algorithms <- c(
  "K-Means + Fisher-Rao",
  "K-Means + Euclidean",
  "Hierarchical + Fisher-Rao",
  "Hierarchical + Euclidean"
)
cols <- c("km_fr", "km_eu", "hc_fr", "hc_eu")

summary_df <- data.frame(
  Algorithm = algorithms,
  Accuracy  = sapply(cols, function(col) mean(results[[col]])),
  SE        = sapply(cols, function(col) sd(results[[col]]) / sqrt(n_rep)),
  row.names = NULL
)
summary_df$Replicated <- sprintf("%.3f ± %.3f", summary_df$Accuracy, summary_df$SE)

knitr::kable(
  summary_df[, c("Algorithm", "Replicated")],
  col.names = c("Algorithm", "Accuracy (mean ± SE)"),
  caption   = "Clustering accuracy over 100 Monte Carlo replications (k=3)"
)
```

---

## Fisher-Rao vs Euclidean: win counts

```{r wins}
km_fr_wins <- sum(results$km_fr > results$km_eu)
hc_fr_wins <- sum(results$hc_fr > results$hc_eu)

cat(sprintf("K-Means  Fisher-Rao wins: %d / %d replications\n", km_fr_wins, n_rep))
cat(sprintf("Hierarch Fisher-Rao wins: %d / %d replications\n", hc_fr_wins, n_rep))
```

Expected pattern (Zhang Ch.8): K-Means Fisher-Rao wins ~98/100; Hierarchical
Fisher-Rao wins ~70/100.

---

## Save results

```{r save}
write.csv(results,    "../results/tables/replication_results.csv", row.names = FALSE)
write.csv(summary_df, "../results/tables/accuracy_summary.csv",    row.names = FALSE)
cat("Results written to results/tables/\n")
```
